---
title: "Rio Grande Minnows"
subtitle: "Zero-inflated CPUE"
author: "Lucas A. Nell"
date: "10/12/2016"
output:
    ioslides_presentation:
        transition: faster
        widescreen: true
runtime: shiny
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
library(magrittr)
library(purrr)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(grid)
library(ggthemes)
library(shiny)
```

```{r input_data, include = FALSE}
if (grepl('theor_ecol$', getwd())){
    setwd('./Pine_dat')
} else if (!grepl('Pine_dat$', getwd())) {
    stop('Unknown directory. It should be in one ending in "theor_ecol" or "Pine_dat".')
}
catch_df <- read_csv('catch by station.csv') %>% 
    mutate(cpue = Count / (Area/100),
           Date = as.Date(paste(Year, Month, '01', sep = '-'), format = '%Y-%m-%d'))
flow_df <- read_csv('ABQ_discharge.csv') %>% 
    select(-ABQ_7000) %>% 
    gather(key = threshold, value = duration, starts_with('ABQ')) %>% 
    mutate(threshold = as.integer(gsub('ABQ_', '', threshold)))
get_flow_dur <- function(yr, thresh = 3000) {
    sapply(yr, 
           function(y){
               flow_df$duration[flow_df$Year == y & flow_df$threshold == thresh]
           })
}
```

```{r gg_theme_adverb, include = FALSE}
# Custom ggplot2 theme
theme_lan <- function(base_size = 10, base_family = 'Helvetica') {
    ggthemes::theme_fivethirtyeight(base_size = base_size, 
                                    base_family = base_family) %+replace%
        theme(
            strip.text = element_text(face = 'bold'),
            axis.title = element_text(size = 11)
        )
}
only_if <- function(condition){
    function(func) {
        if (condition) {
            func
        } else {
            function(., ...) .
        }
    }
}
```



# Sampling design

## Seine captures

```{r sein_pic, fig.align = 'center', out.width = "800px"}
seine_pic <- "https://www.army.mil/e2/c/images/2014/08/18/359126/original.jpg"
knitr::include_graphics(seine_pic)
```


## Sampling effort

```{r area_hist, fig.align = 'center', fig.height = 4, fig.width = 8}
p <- catch_df %>% 
    ggplot(aes(Area)) +
    theme_lan() +
    geom_histogram(bins = 30, fill = 'dodgerblue3') + 
    xlab(expression('Area surveyed (' * m^2 * ')'))

p + scale_x_continuous(labels = scales::comma, breaks = seq(2e3, 6e3, 2e3))
```


#### *Does sampling error increase/decrease with sampling area??*


## Sampling effort (cont.)

```{r locations_line, fig.align = 'center', fig.height = 4, fig.width = 8}
catch_df %>%
    group_by(Date) %>%
    summarize(stations = n()) %>% 
    ggplot(aes(Date, stations)) +
    theme_lan() +
    geom_line(linetype = 1) +
    geom_point(size = 2) +
    scale_x_date() +
    ylab('Number of locations surveyed') + 
    xlab('Year')
```



# Minnow counts

## A schooling species of minnow...

```{r count_distr}
radioButtons("trans", label = "Transformation",
              choices = c('identity', 'log1p', 'log'), selected = 'identity',
             inline = TRUE)
radioButtons("zeros", label = "Remove zeros", choices = c(TRUE, FALSE), selected = FALSE,
             inline = TRUE)
renderPlot({
    catch_df %>%
        only_if(as.logical(input$zeros))(filter)(Count > 0) %>%
        ggplot(aes(cpue)) +
        theme_lan() +
        theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
        geom_histogram(bins = 30, fill = 'dodgerblue3') +
        scale_x_continuous(expression('Minnow counts per' ~ 100 * m^2 ~ 'seined'),
                           trans = input$trans, breaks = c(0, 50, 100, 200),
                           labels = scales::comma)
    }, 
    width = 800)
```


## CPUE by year

```{r count_by_yr, fig.height = 5, fig.width = 8}
count_by_yr <- catch_df %>% 
    ggplot(aes(Date, cpue)) +
    theme_lan() +
    geom_point(aes(color = factor(Year)), 
               position = position_jitter(h = 0, w = 100), alpha = 0.7) +
    stat_summary(fun.data = "mean_cl_boot", size = 0.375, shape = 0, fill = NA) +
    scale_y_continuous(expression('Minnow counts per' ~ 100 * m^2 ~ 'seined'),
                       trans = 'log1p', labels = scales::comma,
                       breaks = c(10, 20, 50, 100, 200)) +
    xlab('Year') +
    theme(legend.position = 'none')
count_by_yr
```



# Water flow

## Flooding {.columns-2}


``` {r flood_pic, fig.align = 'left', out.height = "250px"}
flood_pic <- 'http://www.unm.edu/~cfbyrne/_Media/pulse_02_med_hr.jpeg'
mtn_pic <- 'https://upload.wikimedia.org/wikipedia/en/b/bc/Organ_Mountains.jpg'
spawn_pic <- 
    'https://www.cabq.gov/culturalservices/biopark/aquarium/images/silvery%20minnow2.jpg'
knitr::include_graphics(mtn_pic)
knitr::include_graphics(flood_pic)
```

<br>
<br>
<br>
<br>


``` {r flood_pic2, fig.align = 'right', out.width = "400px"}
knitr::include_graphics(spawn_pic)
```


## Water flow over years

```{r flow_yr, fig.height = 5, fig.width = 8}

flow_p <- flow_df %>% 
    ggplot(aes(Year, duration, fill = factor(threshold))) +
    theme_lan() +
    geom_ribbon(aes(ymax = duration), ymin = 0) +
    ylab("Duration (days)") +
    scale_fill_manual(expression('Flow threshold (' * ft^3 * s^-1 * ')'),
                      values = tail(scales::brewer_pal('seq')(7), -1))
flow_p
```



------

```{r flow_cpue, fig.height = 5, fig.width = 8}
catch_df %>%
    mutate(duration = get_flow_dur(Year)) %>% 
    ggplot(aes(duration, cpue)) +
    theme_lan() +
    geom_point() +
    xlab(expression('Days above 3,000' ~ ft^3 * s^-1)) + 
    scale_y_continuous('CPUE', trans = 'log1p') +
    geom_smooth(method = 'lm', se = FALSE)

lin_mod <- catch_df %>%
    mutate(duration = get_flow_dur(Year)) %>% 
    filter(cpue > 0) %>% 
    lm(cpue ~ duration, data = .)

```


## Hurdle model?

$Y$ = catch per unit effort

$Z$ = indicator variable: $1$ for $Y > 0$, $0$ for $Y = 0$

$E(Y) = Pr(Z = 1) E(Y | Z = 1)$

$E(Y) = \pi \mu$

Then model $\pi$ with logistic regression

And model $\mu$ with linear regression




